Improving the Reliability of the Pavlovian \break Go/No-Go Task

### 2023-07-20 manuscript/main.tex
updated manuscript

\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1,*]{Samuel Zorowitz}
\author[1]{Gili Karni}
\author[2]{Natalie Paredes}
\author[1,3]{Nathaniel Daw}
\author[1,3]{Yael Niv}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychology, University of California, San Diego, USA}
\affil[3]{Department of Psychology, Princeton University, USA}
\affil[*]{Corresponding author (zorowitz@princeton.edu)}

% specify title / suppress date
\title{Improving the Reliability of the Pavlovian Go/No-Go Task}
\date{}

@@

% title page
\maketitle
\thispagestyle{empty}          % remove page number
{\bf Keywords:} tk

@@

%TC:ignore
\abstract{test test test}
\thispagestyle{empty}          % remove page number
%TC:endignore

@@

Humans (and other animals) have an innate tendency to approach rewarding stimuli and shrink from punishing stimuli \cite{carver1994behavioral}. Depending on the context, these hardwired Pavlovian biases can benefit or interfere with instrumental (i.e., action-outcome) learning. This is epitomized in a Pavlovian go/no-go task that orthogonalizes action (Go, No-Go) and outcome valence (reward, punishment) \cite{guitart2012go, guitart2014action}. On the task, human participants are faster to learn actions that are congruent with these Pavlovian response biases (i.e., initiate action to receive reward, inhibit action to avoid punishment) and slower to learn Pavlovian-instrumental incongruent responses (i.e., inhibit action to receive reward, initiate action to avoid punishment). These biases can arise during action selection \cite{guitart2012go} and learning (i.e., increased sensitivity to rewards and punishments following action and inaction, respectively; \cite{swart2017catecholaminergic}).

Individual-differences in performance on the Pavlovian go/no-go task have been the focus of many studies. Pavlovian biases are altered in a number of psychiatric conditions. For example, an increased tendency towards passive avoidance has been observed in individuals with general and social anxiety \cite{mkrtchian2017modeling, peterburs2021impact}, whereas active avoidance is amplified in individuals with a history of suicidal thoughts or behaviors \cite{millner2019suicidal}.  Pavlovian biases are larger in individuals with trauma exposure \cite{ousdal2018impact} and first-episode psychosis \cite{montagnese2020reinforcement}, but attenuated in individuals with depression \cite{huys2016specificity} and schizophrenia \cite{albrecht2016reduction}. Pavlovian biases exhibit a U-shape across the lifespan, decreasing from childhood to young adulthood and increasing again in older age \cite{raab2020adolescents, betts2020learning}. Pavlovian biases have been associated with individual differences in personality (e.g., impulsivity; \cite{eisinger2020pavlovian}) and genetic polymorphisms \cite{richter2014valenced, richter2021motivational}. The expression of Pavlovian biases are also modulated by differences in current state, such as mood \cite{weber2022effects}, anger \cite{wonderlich2020anger}, stress \cite{de2016acute}, and fear \cite{mkrtchian2017threat}.

None of the studies above investigated or reported the reliability of their task performance measures. This is important because the reliability of a measure places an upper bound on the maximum observable correlation between itself and a second measure (e.g., symptom scores; \cite{zorowitz2023improving}). Therefore, as the reliability of a task measure decreases, so too does the power to detect and precisely estimate individual-difference correlations, which in turn increases the possibility of false-negatives \cite{Parsons2019-jw} and bias in correlations reaching statistical significance \cite{gelman2014beyond}. To date, two studies have investigated the psychometrics of the Pavlovian go/no-go task. In a large study of adolescents and young adults, Moutoussis and colleagues \cite{moutoussis2018change} found that both descriptive and model-based measures of performance on the task exhibited poor test-retest reliability. It was unclear, however, if this finding reflected their sample population and/or their lengthy test-retest intervals (6 and 18 months). However, Pike and colleagues \cite{pike2022test} recently reported similar results in an online sample of adults tested twice over over two weeks. Thus, the extant literature indicates performances measures on the Pavlovian go/no-go task are poorly suited for use in individual-differences research.

A second psychometric consideration is task repeatability, or the ability for a task to consistently elicit an expected pattern of behavior across repeated administrations. One common threat to task repeatability is practice effects, where participants’ performance on a task improves with increasing exposure \cite{hausknecht2007retesting, scharfen2018retest}. Practice effects may confound longitudinal studies of task performance, especially studies in which an improvement in performance is hypothesized (e.g., reduction in Pavlovian biases following a psychotherapy intervention \cite{geurts2022aversive}). Moreover, practice effects can harm reliability to the extent they manifest unequally across participants or minimize between-participant variability (i.e., ceiling effects). The repeatability of the Pavlovian go/no-go task has received little attention to date. The only exception is Moutoussis and colleagues \cite{moutoussis2018change} who observed diminished Pavlovian biases in retest performance at 6 and 18 months; however, this finding may reflect expected developmental changes \cite{raab2020adolescents, betts2020learning} rather than practice effects.

The objectives of the current study were to investigate the reliability and repeatability of a gamified version of the Pavlovian go/no-go task \cite{weber2022effects}. Gamification, or the incorporation of (video) game design elements into cognitive tasks, has previously been found to increase participant engagement \cite{sailer2017gamification} and improve the reliability of task measures \cite{kucina2022solution, verdejo2021unified}. Therefore, we hypothesized that our version of the Pavlovian go/no-go task would yield improvements in the test-retest reliability of model-based measures of task performance.

%In Experiment 1, a sample of N=103 participants completed the gamified task four times over one month. We used reinforcement learning models quantify Pavlovian biases in choice behavior. Contrary to our expectations, we observed large practice effects in participants' performance, including the virtual disappearance of Pavlovian biases, that in turn impacted the reliability of estimated model parameters. In Experiment 2, a sample of N=110 participants completed a modified version of the Pavlovian go/no-go task --- one that prevents participants from using alternative strategies to complete the task --- three times over two weeks. Here we observed comparably smaller practice effects and, crucially, acceptable reliability of estimated model parameters.

@@

A total of N=148 participants were recruited in May, 2020 from Amazon Mechanical Turk via CloudResearch \cite{litman2017turkprime} to participate in an online behavioral experiment. Participants were eligible to participate if they were at least 18 years old and resided in the United States. Following best practice recommendations \cite{robinson2019tapped}, no other inclusion criteria were applied. This study was approved by the Institutional Review Board of Princeton University (\#5291) and all participants provided informed consent. Total study duration was 15--20 minutes. Participants received monetary compensation for their time (rate: USD \$12/hr), plus an incentive-compatible bonus up to \$1.50 based on task performance.

Data from N=45 participants who completed the first session were excluded prior to analysis (see ``Exclusion criteria'' in the Supplementary Materials), leaving a final sample of N=103 participants. These participants were re-invited to complete a follow-up experiment 3, 14, and 28 days later. Once invited, participants were permitted 48 hours to complete the follow-up experiment. Retention was high for each follow-up session (Day 3: N=94 [91.3\%]; Day 14: N=92 [89.3\%]; Day 28: N=89 [86.4\%]). In addition to the performance bonus, participants received a retention bonus of \$1.00 for each completed follow-up session. Detailed demographic information is presented in Table \ref{tab:demographics}. The majority of participants identified as men (55 men; 47 women; 1 non-binary) and were 35.5 years old on average (range: 20--69 years).

@@

The task involved four trial types that differed in their correct action (Go, No-Go) and outcome domain (reward, punishment; Figure \ref{fig:task_schematic}B). Specifically, the four trial types were: go to win points (GW); no-go to win points (NGW); go to avoid losing points (GAL); and no-go to avoid losing points (NGAL). Note that GW and NGAL trials are Pavlovian-instrumental `congruent', because there is a match between the correct response and the expected approach/avoidance bias for each, whereas NGW and GAL trials are Pavlovian-instrumental `incongruent'. For the rewarding trials (GW, NGW), the possible outcomes were +10 or +1 points; for the punishing trials (GAL, NGAL), they were -1 or -10 points. The outcome domain of each robot was explicitly signaled to participants by a blue or orange `scanner light' (randomized within participants across sessions). Outcomes were probabilistic such that participants received the better of the two possible outcomes with 80\% chance if they made the correct action, and received the worse of the possible outcomes with 80\% chance if they made the incorrect action (Figure \ref{fig:task_schematic}C).

@@

@@

We first evaluated participants' choice behavior using five performance indices: overall percent correct responses; go bias, or the difference in correct responses between go and no-go trials; outcome bias, or the difference in correct responses between rewarding and punishing trials; Pavlovian bias, or the difference in correct responses between Pavlovian-instrumental congruent and incongruent trials; and feedback sensitivity, or the difference in correct responses between trials following veridical or sham feedback. For each session and index, we tested if the median value across participants was significantly different than zero (or 50\% for overall percent correct responses). We use the median due to skew in the performance indices. We also tested if the median value of each index was significantly different between each pair of sessions. P-values were derived via permutation testing, where a null distribution of values were obtained by permuting the condition labels (for within-session tests) or session labels (for between-session tests) 5000 times. Within-session tests were not corrected for multiple comparisons as each test constitutes an individual hypothesis test; between-session tests were corrected for multiple comparisons using the familywise error rate correction \cite{winkler2014permutation}.

@@

We fit a set of seven nested reinforcement learning (RL) models to participants' choice data. All models are variants of the Rescorla-Wagner model and have previously been used to predict choice behavior on this task \cite{guitart2012go, mkrtchian2017modeling, moutoussis2018change, swart2017catecholaminergic}. In the base model (M1), the probability a participant chooses an action in response to stimulus $k$ is determined by the difference in stimulus-action values ($Q_k(\text{Go}) - Q_k(\text{No-Go}))$, scaled an outcome sensitivity parameter ($\beta$), and passed through the logistic function. In turn, stimulus-action values are learned via the Rescorla-Wagner update rule as governed by a learning rate ($\eta$). Model 2 is augmented with a single static approach/avoidance parameter ($\tau$) that biases action towards Go responses if $\tau_j > 0$ and No-Go responses if $\tau < 0$. The succeeding model (M3) has independent approach/avoidance parameters per outcome domain ($\tau_+$, $\tau_-$). Models 4 and 5 have either independent outcome sensitivity ($\beta_+$, $\beta_-$; M4) or learning rate parameters ($\eta_+$, $\eta_-$; M5) per outcome domain. In Model 6, all three parameter types are independent across outcome domains. Finally, the most complex model (M7) includes a lapse rate ($\xi$). Complete model equations and descriptions are provided in the Supplementary Materials.

@@

The fits of the models to the data were assessed using posterior predictive checks. Specifically, we inspected each model's ability to reproduce both group-averaged learning curves (by trial type) and each participant's proportion of go responses (by trial type). Model fits were compared using Bayesian leave-one out cross-validation (LOO; \cite{vehtari2017practical}). A credible improvement in model fit was defined as a difference in LOO values four times larger than its corresponding standard error \cite{Vehtari_undated-tc}.

We investigated the reliability of the model parameters for the best-fitting model also using a Bayesian hierarchical modeling framework \cite{rouder2019psychometrics}. Briefly, the best-fitting model was fit to participants' choice data with partial-pooling over blocks within a session or over whole datasets between sessions. Split-half and test-retest reliability was respectively calculated as the Pearson correlation between model parameters estimated from each block and each pair of sessions. Complete details of the estimation procedure are reported in the Supplementary Materials.

@@

%TC:ignore
\begin{figure}[hpt]
    \centerline{\includesvg[inkscapelatex=false,width=1.00\textwidth]{figures/fig02.svg}}
    \caption{Summary of behavior in Experiment 1. (A) Group-averaged learning curves for each trial type and session. Shaded regions indicate 95\% bootstrapped confidence intervals. (B) Group-averaged performance for each session. Performance indices from left-to-right: Correct responses, or overall accuracy; Go bias, or difference in accuracy between Go and No-Go trials; Congruency effect, or difference in accuracy between Pavlovian congruent (GW, NGAL) and incongruent (NGW, GAL) trials; and Feedback sensitivity, or the difference in accuracy on trials following veridical and sham feedback. Blue shading indicates significant differences in comparison to all other sessions. (C) The percentage of participants, for each session and trial type, exhibiting at- or below-chance performance ($< 60\%$ response accuracy; grey), moderate performance ($\geq 60\%$ response accuracy; light blue), or near-perfect performance ($\geq 90\%$ response accuracy; dark blue).}
    \label{fig:exp01_behavior}
\end{figure}
%TC:endignore

Trial-by-trial choice behavior for each session is presented in Figure \ref{fig:exp01_behavior}A. Performance in the first session qualitatively conformed to the expected pattern of results, i.e., worse performance on Pavlovian-instrumental incongruent trials. However, this effect was ostensibly diminished in all follow-up sessions. We turn to the descriptive analyses to formally test these trends.

The group-averaged performance indices per session are summarized in Figure \ref{fig:exp01_behavior}B. On Day 0, the overall percent correct responses was 85.0\%. Performance improved in all subsequent sessions. Pairwise comparisons confirmed that performance was indeed worse on Day 0 compared to each follow-up session (all $p < 0.001$); no other comparisons were significant. Participants' self-reported mood and anxiety were largely stable over the same period (Figure \ref{fig:figS05}), indicating this shift in performance more likely reflects practice effects rather than changes in participants' state.

Across sessions, participants made more correct responses on go trials than no-go trials. The go bias was significantly reduced in all follow-up sessions compared to Day 0 (all $p < 0.001$); so too was the bias on Day 28 compared to Day 3 ($p < 0.001$). Similarly, participants made more correct responses on Pavlovian-congruent than incongruent trials. Again, the Pavlovian bias was significantly reduced in all follow-up sessions compared to Day 0 (all $p < 0.001$); no other comparisons were significant. Participants made marginally more correct responses on rewarding trials on Day 3 only.

A similar pattern of results was observed for feedback sensitivity. Across sessions, participants made more correct responses following veridical compared to sham feedback (Day 0: $\Delta$M = 0.094, $d = 1.250$, $p < 0.001$, 95\% CI = [0.072, 0.115]; Day 3: $\Delta$M = 0.045, $d = 0.770$, $p < 0.001$, 95\% CI = [0.034, 0.062]; Day 14: $\Delta$M = 0.030, $d = 0.581$, $p < 0.001$, 95\% CI = [0.018, 0.047]; Day 28: $\Delta$M = 0.028, $d = 0.704$, $p < 0.001$, 95\% CI = [0.022, 0.043]). Feedback sensitivity was significantly less in all follow-up sessions compared to Day 0 (all $p < 0.001$); no other comparisons were significant.

The results so far summarize group-averaged performance but do not provide much insight into individual variability. As such, the proportion of participants who exhibited chance-level ($<$60\% correct responses), intermediate ($\geq$60\%), or near-ceiling performance ($\geq$90\%) by session and trial type is presented in Figure \ref{fig:exp01_behavior}C. Excepting GW trials, the percentage of participants nearing ceiling-level performance increases from the minority on Day 0 to the majority in all follow-up sessions. Two-way chi-squared tests confirm this trend (GW: $\chi^2 (6) = 8.149$, $p = 0.227$; NGW: $\chi^2 (6) = 55.458$, $p < 0.001$; GAL: $\chi^2 (6) = 42.191$, $p < 0.001$; NGAL: $\chi^2 (6) = 39.287$, $p < 0.001$). In sum, the increases in task performance (and accompanying reductions in choice biases) with repeated testing observed at the group-level extends to the majority of participants.

@@

%TC:ignore
\begin{table}[b!]
    \centering
    \begin{tabular}{clccr}
        \toprule
        Model & Parameters & Acc. (\%) & LOO & $\Delta$ LOO (se) \\
        \midrule
        1 & $\beta, \eta$ & 87.5 & -151457.9 & -5602.6 (68.3) \\
        2 & $\beta, \tau, \eta$ & 89.0 & -154011.9 & -3048.6 (51.2) \\
        3 & $\beta, \tau_+, \tau_-, \eta$ & 89.8 & -155817.8 & -1242.7 (31.3) \\
        4 & $\beta_+, \beta_-, \tau_+, \tau_-, \eta$ & 89.8 & -156261.6 & -798.8 (22.6) \\
        5 & $\beta, \tau_+, \tau_-, \eta_+, \eta_-$ & 89.9 & -156265.9 & -794.6 (20.7) \\
        6 & $\beta_+, \beta_- \tau_+, \tau_-, \eta_+, \eta_-$ & 89.9 & -156401.8 & -658.6 (18.8) \\
        7 & $\beta_+, \beta_- \tau_+, \tau_-, \eta_+, \eta_-, \xi$ & 90.1 & -157060.5 & \multicolumn{1}{c}{-} \\
        \bottomrule
\end{tabular}
    \caption{Model comparison collapsing across sessions. Notes: Acc. = trial-level predictive accuracy between observed and model-predicted choice. LOO = Bayesian leave-one-out cross-validation scores presented in deviance scale (i.e., smaller numbers indicate better fit). $\Delta$LOO = difference in LOO values between each model and the best-fitting model (M7).}
    \label{tab:exp1_mc_abbr}
\end{table}
%TC:endignore

@@

>>>>

\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1,*]{Samuel Zorowitz}
\author[1]{Gili Karni}
\author[2]{Natalie Paredes}
\author[1,3]{Nathaniel Daw}
\author[1,3]{Yael Niv}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychology, University of California, San Diego, USA}
\affil[3]{Department of Psychology, Princeton University, USA}
\affil[*]{Corresponding author (szorowi1@gmail.com)}

% specify title / suppress date
\title{Improving the Reliability of the Pavlovian \break Go/No-Go Task}
\date{}

@@

% title page
\maketitle
\thispagestyle{empty}          % remove page number
{\bf Keywords:} Pavlovian go/no-go task; Pavlovian bias; reinforcement learning; reliability

@@

%TC:ignore
\abstract{

\noindent \textbf{Background:} The Pavlovian go/no-go task is commonly used to measure individual differences in Pavlovian biases and their interaction with instrumental learning. However, prior research has found suboptimal reliability in computational model-based performance measures for this task, limiting its usefulness in individual-differences research. These studies did not make use of several strategies that previously shown to enhance task-measure reliability (e.g., task gamification, hierarchical Bayesian modeling for model estimation). Here we investigated if such approaches could improve the task's reliability.\\

\noindent \textbf{Methods:} Across two experiments, we recruited two independent samples of adult participants (N=103, N=110) to complete two novel, gamified versions of the Pavlovian go/no-go task multiple times over several weeks. We used hierarchical Bayesian modeling to derive reinforcement learning model-based indices of participants' task performance, and additionally to estimate the reliability of these measures.\\

\noindent \textbf{Results:} In Experiment 1, we observed considerable and unexpected practice effects, with most participants reaching near-ceiling levels of performance with repeat testing. Consequently, the test-retest reliability of some model parameters was unacceptable (range: 0.379--0.973). In Experiment 2, participants completed a modified version of the task designed to lessen these practice effects. We observed greatly reduced practice effects and improved estimates of the test-retest reliability (range: 0.696--0.989).\\

\noindent \textbf{Conclusion:} The results demonstrate that model-based measures of performance on the Pavlovian go/no-go task can reach levels of reliability sufficient for use in individual-differences research. However, additional investigation is necessary to validate the modified version of the task in other populations and settings.
}
\thispagestyle{empty}          % remove page number
%TC:endignore

@@

Humans (and other animals) have an innate tendency to approach rewarding stimuli and shrink from punishing stimuli \cite{carver1994behavioral}. Depending on the context, these hardwired Pavlovian biases can either benefit or interfere with instrumental (i.e., action-outcome) learning. This is epitomized in the Pavlovian go/no-go task in which the required action (Go, No-Go) and outcome valence (reward, punishment) are orthogonalized \cite{guitart2012go, guitart2014action}. In the task, participants are typically faster to learn actions that are congruent with Pavlovian response biases (i.e., initiate action to receive reward, inhibit action to avoid punishment) and slower to learn Pavlovian-instrumental incongruent responses (i.e., inhibit action to receive reward, initiate action to avoid punishment). %These biases can arise during action selection \cite{guitart2012go} and learning (i.e., increased sensitivity to rewards and punishments following action and inaction, respectively; \cite{swart2017catecholaminergic}).

The Pavlovian go/no-go task has been used in innumerable studies to probe individual-differences in reward and punishment learning, of which many have reported changes in Pavlovian biases as a function of psychiatric conditions. For example, an increased tendency towards passive avoidance has been observed in individuals with general and social anxiety \cite{mkrtchian2017modeling, peterburs2021impact}, whereas active avoidance is amplified in individuals with a history of suicidal thoughts or behaviors \cite{millner2019suicidal}. Pavlovian biases are larger in individuals with trauma exposure \cite{ousdal2018impact} and first-episode psychosis \cite{montagnese2020reinforcement}, but attenuated in individuals with depression \cite{huys2016specificity} and schizophrenia \cite{albrecht2016reduction}. Pavlovian biases have also been associated with individual differences in personality (e.g., impulsivity; \cite{eisinger2020pavlovian}) and genetics \cite{richter2014valenced, richter2021motivational}. In developmental and lifespan research, Pavlovian biases have been shown to exhibit a U-shape, decreasing from childhood to young adulthood and increasing again in older age \cite{raab2020adolescents, betts2020learning}. At a finer temporal scale, Pavlovian biases are also reportedly modulated by state effects including mood \cite{weber2022effects}, anger \cite{wonderlich2020anger}, stress \cite{de2016acute}, and fear \cite{mkrtchian2017threat}.

Despite its extensive use, the Pavlovian go/no-go task has received less psychometric investigation. What studies exist suggest the task, in its canonical form, is suboptimal for use in individual-differences correlational research. Specifically, two independent studies found that descriptive and computational model-based measures of performance on the Pavlovian go/no-go task exhibited suboptiomal test-retest reliability over short (two-week) and long (6-, 18-month) retest intervals \cite{moutoussis2018change, pike2022test}. This is important because the reliability of a measure places an upper bound on the maximum observable correlation between itself and a second measure (e.g., symptom scores; \cite{Spearman1904-mo}). Therefore, as reliability decreases, so too does statistical power; in turn, this increases the possibility of false-negatives \cite{Parsons2019-jw} and bias in correlations that do reach statistical significance \cite{gelman2014beyond}. One of these studies also reported evidence of practice effects (i.e., improved performance and diminished Pavlovian biases with repeated testing; \cite{moutoussis2018change}). Practice effects are of additional concern for individual-differences research because they can imperil reliability (e.g., by minimizing between-participants variability) and obscure effects of interest (e.g., changes in task performance following changes in mood).

These previous studies of the psychometrics of the Pavlovian go/no-go task, however, did not make use of multiple strategies for improving the reliability of cognitive task measures \cite{zorowitz2023improving}. For example, prior research has found that gamification, or the incorporation of (video) game design elements into cognitive tasks, can promote participant engagement \cite{sailer2017gamification} and improve the reliability of task measures \cite{kucina2022solution, verdejo2021unified}. Moreover, the use of hierarchical Bayesian models --- which exert a pooling effect on person-level variables, in effect correcting them for measurement error \cite{haines2023classical, rouder2019psychometrics} --- have been frequently shown to improve the reliability of task measures \cite{sullivan2022enhancing, brown2020improving, waltmann2022sufficient}. Finally, practice effects can be lessened by designing tasks in such a way that prevents participants from discovering and using task-specific knowledge to enhance their performance on subsequent attempts \cite{mclean2018towards}.

The aim of the current study was to investigate the reliability and repeatability of a novel version of the Pavlovian go/no-go task. We conducted two experiments involving two independent samples of adult participants who completed a gamified version of the task multiple times over several weeks. We used hierarchical Bayesian models to derive reinforcement learning model-based indices of their task performance, and additionally to estimate the reliability of these measures. In Experiment 1, participants unexpectedly exhibited large practice effects, which negatively impacted the test-retest reliability of the performance measures. To address this issue, in Experiment 2, participants completed a modified version of the task that reduced practice effects, and led to significant improvements in the test-retest reliability of the reinforcement learning model parameters.

@@

A total of N=148 participants were recruited in May, 2020 from Amazon Mechanical Turk via CloudResearch \cite{litman2017turkprime} to participate in an online behavioral experiment. Participants were eligible to participate if they were at least 18 years old and resided in the United States. Following best practice recommendations \cite{robinson2019tapped}, no other inclusion criteria were applied. This study was approved by the Institutional Review Board of Princeton University and all participants provided informed consent. Total study duration was 15--20 minutes. Participants received monetary compensation for their time (rate: USD \$12/hr), plus an incentive-compatible bonus up to \$1.50 based on task performance.

Data from N=45 participants who completed the first session were excluded prior to analysis (see ``Exclusion criteria'' below), leaving a final sample of N=103 participants. These participants were re-invited to complete a follow-up experiment 3, 14, and 28 days later. Once invited, participants were permitted 48 hours to complete the follow-up experiment. Retention was high for each follow-up session (Day 3: N=94 [91.3\%]; Day 14: N=92 [89.3\%]; Day 28: N=89 [86.4\%]). In addition to the performance bonus, participants received a retention bonus of \$1.00 for each completed follow-up session. Detailed demographic information is presented in Table \ref{tab:demographics}. The majority of participants identified as men (55 men; 47 women; 1 non-binary) and were 35.5 years old on average (range: 20--69 years).

@@

The task involved four trial types that differed by their correct action (Go, No-Go) and outcome domain (reward, punishment; Figure \ref{fig:task_schematic}B). Specifically, the four trial types were: go to win points (GW); no-go to win points (NGW); go to avoid losing points (GAL); and no-go to avoid losing points (NGAL). Note that GW and NGAL trials are Pavlovian-instrumental `congruent' because there is a match between the correct response and the expected approach/avoidance bias for each. In contrast, NGW and GAL trials are Pavlovian-instrumental `incongruent'. For the rewarding trials (GW, NGW), the possible outcomes were +10 or +1 points; for the punishing trials (GAL, NGAL), they were -1 or -10 points. The outcome domain of each robot was explicitly signaled to participants by a blue or orange `scanner light' (randomized within participants across sessions). Outcomes were probabilistic such that participants received the better of the two possible outcomes with 80\% chance if they made the correct action, and received the worse of the possible outcomes with 80\% chance if they made the incorrect action (Figure \ref{fig:task_schematic}C).

@@

\subsubsection*{Exclusion criteria}

To ensure data quality, the data from multiple participants from the initial session were excluded prior to analysis. In Experiment 1, the data from multiple participants who completed the experiment on Day 0 were excluded prior to analysis for one or both of the following reasons: failing more than one attention check embedded in the self-report measures (N=13; \cite{zorowitz2023inattentive}) or exhibiting chance-level performance ($<$55\% correct responses) on go to win trials (N=43). Data from N=45 participants who completed the first session were excluded based on these criteria, leaving a final sample of N=103 participants. No exclusions were applied to subsequent session data.

@@

We first evaluated participants' choice behavior using five performance measures: overall percent correct responses; go bias, or the difference in correct responses between Go and No-Go trials; outcome bias, or the difference in correct responses between rewarding and punishing trials; Pavlovian bias, or the difference in correct responses between Pavlovian-instrumental congruent and incongruent trials; and feedback sensitivity, or the difference in correct responses between trials following veridical or sham feedback. For each session and measure, we tested if the median value across participants was significantly different than zero (or 50\% for overall percent correct responses). We used the median due to skew in the performance measures. We also tested if the median value of each measure was significantly different between each pair of sessions. P-values were derived via permutation testing, where a null distribution of values were obtained by permuting the condition labels (for within-session tests) or session labels (for between-session tests) 5,000 times. Within-session tests were not corrected for multiple comparisons as each test constitutes an individual hypothesis test; however, between-session tests were corrected, using the family-wise error rate correction \cite{winkler2014permutation}, because they constitute a disjunctive test \cite{rubin2021adjust}.

@@

We fit a set of seven nested reinforcement learning (RL) models to participants' choice data. All models are variants of the Rescorla-Wagner model and have previously been used to predict choice behavior on this task \cite{guitart2012go, mkrtchian2017modeling, moutoussis2018change, swart2017catecholaminergic}. In the base model (M1), the probability a participant chooses an action in response to stimulus $k$ is determined by the difference in stimulus-action values ($Q_k(\text{Go}) - Q_k(\text{No-Go}))$, scaled an outcome sensitivity parameter ($\beta$), and passed through the logistic function. In turn, stimulus-action values are learned via the Rescorla-Wagner update rule as governed by a learning rate ($\eta$). Model 2 is augmented with a single static approach/avoidance parameter ($\tau$) that biases action towards Go responses if $\tau_j > 0$ and No-Go responses if $\tau < 0$. Model 3 (M3) introduces independent approach/avoidance parameters per outcome domain ($\tau_+$, $\tau_-$). Models 4 and 5 have either independent outcome sensitivity ($\beta_+$, $\beta_-$; M4) or learning rate parameters ($\eta_+$, $\eta_-$; M5) per outcome domain. In Model 6, all three parameter types ($\beta$, $\tau$, $\eta$) are independent across outcome domains. Finally, the most complex model (M7) includes a lapse rate ($\xi$). Complete model equations and descriptions are provided in the Supplementary Materials.

@@

The fits of the models to the data were assessed using posterior predictive checks. Specifically, we inspected each model's ability to reproduce both group-averaged learning curves (by trial type) and each participant's proportion of go responses (by trial type). Model fits were compared using approximate leave-one-out cross-validation via Pareto smoothed importance sampling (PSIS-LOO; \cite{vehtari2017practical}). A credible improvement in model fit was defined as a difference in PSIS-LOO values four times larger than its corresponding standard error \cite{Vehtari_undated-tc}.

We investigated the reliability of the model parameters for the best-fitting model also using a Bayesian hierarchical modeling framework \cite{rouder2019psychometrics}. Briefly, the best-fitting model was fit to participants' choice data with partial pooling over blocks within a session (for split-half reliability) or over whole datasets between sessions (for test-retest reliability). Complete details of the estimation procedure are reported in the Supplementary Materials. Split-half and test-retest reliability was then respectively calculated as the Pearson correlation between model parameters estimated from each block and each pair of sessions \cite{brown2020improving, pike2022test}. We use the Pearson correlation because we are primarily interested in the consistency of rank-ordering of participants' parameter estimates over time. Though arbitrary, we follow convention and define $\rho \geq 0.7$ as the threshold for ``acceptable'' reliability \cite{cicchetti1994guidelines}.

@@

%TC:ignore
\begin{figure}[hpt]
    \centerline{\includesvg[inkscapelatex=false,width=1.00\textwidth]{figures/fig02.svg}}
    \caption{Participants in Experiment 1 show large practice effects on the standard Pavlovian go/no-go task. (A) Group-averaged learning curves for each trial type and session. Shaded regions indicate 95\% bootstrapped confidence intervals. (B) Group-averaged performance for each session. Performance measures from left-to-right: Correct responses, or overall accuracy; Go bias, or difference in accuracy between Go and No-Go trials; Congruency effect, or difference in accuracy between Pavlovian congruent (GW, NGAL) and incongruent (NGW, GAL) trials; and Feedback sensitivity, or the difference in accuracy on trials following veridical and sham feedback. **Denotes significant pairwise difference ($p < 0.05$, corrected for multiple comparisons). (C) The percentage of participants, for each session and trial type, exhibiting at- or below-chance performance ($< 60\%$ response accuracy; grey), intermediate performance ($\geq 60\%$ response accuracy; light blue), or near-perfect performance ($\geq 90\%$ response accuracy; dark blue).}
    \label{fig:exp01_behavior}
\end{figure}
%TC:endignore

Trial-by-trial choice behavior for each session is presented in Figure \ref{fig:exp01_behavior}A. Performance in the first session qualitatively conformed to the expected pattern of results (i.e., worse performance on Pavlovian-instrumental incongruent trials). However, this effect was seemingly diminished in all follow-up sessions. We turn to the descriptive analyses to formally test these trends.

The group-averaged performance measures by session are summarized in Figure \ref{fig:exp01_behavior}B. (Complete descriptive statistics are reported in Table~\ref{tab:stats_exp01}.) Participants made the correct response on 85.0\% of trials on Day 0, which increased to near-ceiling levels in all subsequent sessions. Pairwise comparisons confirmed that performance was indeed worse on Day 0 compared to each follow-up session (all $p < 0.001$); no other comparisons were significant. Participants' self-reported mood and anxiety were largely stable over the same period (Figure \ref{fig:figS05}), indicating this shift in performance more likely reflects practice effects rather than changes in participants' state.

Across sessions, participants made more correct responses on Go trials than No-Go trials. However, the go bias was significantly reduced in all follow-up sessions compared to Day 0 (all $p < 0.001$); so too was it on Day 28 compared to Day 3 ($p < 0.001$). Similarly, participants made more correct responses on Pavlovian-congruent than incongruent trials. As with the Go bias, the Pavlovian bias was significantly reduced in all follow-up sessions compared to Day 0 (all $p < 0.001$); no other comparisons were significant. Regarding the valence bias, participants made marginally but significantly more correct responses on rewarding trials on Day 3 only; no significant difference were observed between sessions.

A similar pattern of results was observed for feedback sensitivity. Across sessions, participants made more correct responses following veridical compared to sham feedback (all $p < 0.001$). However, feedback sensitivity was significantly reduced in all follow-up sessions compared to Day 0 (all $p < 0.001$); no other comparisons were significant. This is consistent with participants' learning curves which show, excluding Day 0, participants quickly learn the correct action for each stimulus and maintain this policy despite 20\% sham feedback.

The results so far summarize group-averaged performance but do not provide much insight into individual differences. As such, the proportion of participants who exhibited chance-level ($<$60\% correct responses), intermediate ($\geq$60\%), or near-ceiling performance ($\geq$90\%) by session and trial type is presented in Figure \ref{fig:exp01_behavior}C. Excepting GW trials, the percentage of participants nearing ceiling-level performance increases from the minority on Day 0 to the majority in all follow-up sessions. Two-way chi-squared tests confirm this trend (GW: $\chi^2 (6) = 8.149$, $p = 0.227$; NGW: $\chi^2 (6) = 55.458$, $p < 0.001$; GAL: $\chi^2 (6) = 42.191$, $p < 0.001$; NGAL: $\chi^2 (6) = 39.287$, $p < 0.001$). In sum, the improvements in task performance (and accompanying reductions in choice biases) with repeat testing observed at the group-level extends to the majority of participants.

@@

%TC:ignore
\begin{table}[b!]
    \centering
    \begin{tabular}{clccr}
        \toprule
        Model & Parameters & Acc. (\%) & PSIS-LOO & $\Delta$ PSIS-LOO (se) \\
        \midrule
        1 & $\beta, \eta$ & 87.5 & -151457.9 & -5602.6 (68.3) \\
        2 & $\beta, \tau, \eta$ & 89.0 & -154011.9 & -3048.6 (51.2) \\
        3 & $\beta, \tau_+, \tau_-, \eta$ & 89.8 & -155817.8 & -1242.7 (31.3) \\
        4 & $\beta_+, \beta_-, \tau_+, \tau_-, \eta$ & 89.8 & -156261.6 & -798.8 (22.6) \\
        5 & $\beta, \tau_+, \tau_-, \eta_+, \eta_-$ & 89.9 & -156265.9 & -794.6 (20.7) \\
        6 & $\beta_+, \beta_- \tau_+, \tau_-, \eta_+, \eta_-$ & 89.9 & -156401.8 & -658.6 (18.8) \\
        7 & $\beta_+, \beta_- \tau_+, \tau_-, \eta_+, \eta_-, \xi$ & 90.1 & -157060.5 & \multicolumn{1}{c}{-} \\
        \bottomrule
\end{tabular}
    \caption{Model comparison collapsing across sessions. Notes: Acc. = trial-level choice prediction accuracy between observed and model-predicted Go responses. PSIS-LOO = approximate leave-one-out cross-validation scores presented in deviance scale (i.e., smaller numbers indicate better fit). \break $\Delta$PSIS-LOO = difference in PSIS-LOO values between each model and the best-fitting model (M7).}
    \label{tab:exp1_mc_abbr}
\end{table}
%TC:endignore

@@

\subsubsection*{Exclusion criteria}
